Based on polling as of 27 October 2024, Harris expects to win on 49.2%, leads Trump by 2% in 2024 US election

### 2024-10-30 paper/paper.qmd
update paper

We conducted a regression analysis to predict support for Kamala Harris
and Donald Trump in the 2024 U.S. presidential election, including
models for Trump for comparison. Our analysis includes four models:
**unweighted linear models** and **weighted linear models** for both
candidates. These models predict each candidate’s support as a
continuous outcome across different states, using **pollster**,
**national poll**, and **population** as key predictors. The weighted
models additionally incorporate adjustments for recency, sample size,
and poll quality to account for potential biases. This setup aligns with
the New York Times methodology for election polling, which assigns
weights based on poll quality, sample size, and recency to better
reflect voter sentiment [@nytimes].

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 1000 
  ),
  pollster = factor("YouGov", levels = levels(harris_data$pollster)),  # Ensure it matches training levels
  national_poll = rep(c(0, 1), length(1000)),  # Create a sequence for national_poll from 0 to 1
  population = factor("lv", levels = levels(harris_data$population)),  # Use the mean population or any relevant value
  sample_size = seq(min(0),max(2300),length.out=1000)
)|>
  mutate(
    # Sample size weight (assuming a maximum of 2300 responses)
    sample_size_weight = pmin(sample_size / 2300, 1),  # Set sample size; adjust accordingly

    # Pollster quality weight (assuming you want to use a dummy numeric grade, e.g., 2)
    pollster_quality_weight = 2 / 4,  # Example; replace with actual numeric grade if available

    # Recency weight (this can be based on the date difference from a certain cutoff, e.g., "2024-11-05")
    recency = as.numeric(difftime(as.Date("2024-11-05"), as.Date("2024-08-01") + (1:1000), units = "days")),  # Adjust for the sequence
    recency_weight = exp(-recency * 0.1),  # Adjust decay rate as needed

    # Poll frequency weight (you can set a condition based on a hypothetical recent poll count)
    poll_frequency_weight = ifelse(national_poll > 0.5, 1, 0.5),  # Example logic for poll frequency

    # Combine all weights into a final combined weight
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight
  )

new_data_trump <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 1000
  ),
  pollster = factor("YouGov", levels = levels(harris_data$pollster)),  # Ensure it matches training levels
  national_poll = rep(c(0, 1), length(1000)),  # Create a sequence for national_poll from 0 to 1
  population = factor("lv", levels = levels(harris_data$population)),  # Use the mean population or any relevant value
  sample_size = seq(min(0),max(2300),length.out=1000)
)|>
  mutate(
    # Sample size weight (assuming a maximum of 2300 responses)
    sample_size_weight = pmin(sample_size / 2300, 1),  # Set sample size; adjust accordingly

    # Pollster quality weight (assuming you want to use a dummy numeric grade, e.g., 2)
    pollster_quality_weight = 2 / 4,  # Example; replace with actual numeric grade if available

    # Recency weight (this can be based on the date difference from a certain cutoff, e.g., "2024-11-05")
    recency = as.numeric(difftime(as.Date("2024-11-05"), as.Date("2024-08-01") + (1:1000), units = "days")),  # Adjust for the sequence
    recency_weight = exp(-recency * 0.1),  # Adjust decay rate as needed

    # Poll frequency weight (you can set a condition based on a hypothetical recent poll count)
    poll_frequency_weight = ifelse(national_poll > 0.5, 1, 0.5),  # Example logic for poll frequency

    # Combine all weights into a final combined weight
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-support-predicted
#| fig-cap: Predicted Support Trends for Kamala Harris and Donald Trump using fitted model
# Predictions for Harris
new_data_harris <- new_data_harris |>
  mutate(
    predicted_unweighted = predict(model_harris_unweighted, newdata = new_data_harris),
    predicted_weighted = predict(model_harris_weighted, newdata = new_data_harris)
  )

# Predictions for Trump
new_data_trump <- new_data_trump |>
  mutate(
    predicted_unweighted = predict(model_trump_unweighted, newdata = new_data_trump),
    predicted_weighted = predict(model_trump_weighted, newdata = new_data_trump)
  )

# Plot the predictions
library(dplyr)
library(ggplot2)

# Define the cutoff date for recent data
recent_date_cutoff <- as.Date("2024-05-01")

# Filter combined_predictions to include only data from the recent date cutoff onward
combined_predictions_filtered <- combined_predictions %>%
  filter(end_date >= recent_date_cutoff)

# Determine the maximum end date for setting the x-axis limit
max_end_date <- max(combined_predictions_filtered$end_date)

# Plotting with the filtered data
ggplot(combined_predictions_filtered, aes(x = end_date, color = candidate)) +
  geom_smooth(aes(y = predicted_unweighted, linetype = "Unweighted"), 
              method = "loess", se = FALSE, size = 1) +  # Unweighted model with smoothing
  geom_smooth(aes(y = predicted_weighted, linetype = "Weighted"), 
              method = "loess", se = FALSE, size = 1) +  # Weighted model with smoothing
  labs(title = "Predicted Support for Kamala Harris and Donald Trump in 2024 US Election \n(From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Weights: recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight\nLinear Regression Model: candidate support percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_linetype_manual(name = "Linear Model Type", values = c("Unweighted" = "dashed", "Weighted" = "solid")) +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

```
The predicted support for Kamala Harris and Donald Trump, as derived from multiple linear regression models, is displayed in the graph above. The analysis spans from August 1, 2024, to October 27, 2024, and highlights both unweighted and weighted predictions for each candidate.

For Kamala Harris, the unweighted model predicts her support to remain relatively stable, maintaining an approximate support level of 49%. This trend is visually represented by the blue dashed line, which indicates minimal fluctuations over time. Conversely, the weighted model shows a more pronounced trend, demonstrating a gradual increase in predicted support, suggesting a potential shift in voter sentiment towards Harris as the election date approaches.

In contrast, Donald Trump's predicted support, represented by the red lines, exhibits a downward trend in both unweighted and weighted models. The unweighted model indicates that his support is expected to hover around 48%, while the weighted model predicts a slight decline to approximately 46%. This trend may reflect challenges in maintaining voter enthusiasm in the face of changing public sentiment.

At the end of the polling period, we predict that Kamala Harris leads Donald Trump by 2% in the unweighted model (48% vs. 46%) and by about 2% in the weighted model (49.1% vs. 47.3%).

@@

# Additional data details

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
# Load necessary libraries
library(boot)    # For bootstrapping
library(caret)   # For cross-validation

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Diagnostics of model using residual vs fitted plot and norm Q-Q plot\-Support for Harris"
#| fig-subcap: ["Residual Plot for Unweighted Model", "Q-Q Plot for Unweighted Model", "Residual Plot for Weighted Model", "Q-Q Plot for Weighted Model"]
#| layout-ncol: 2

>>>>

We conducted a regression analysis to predict support for Kamala Harris
and Donald Trump in the 2024 U.S. presidential election, including
models for Trump for comparison. Our analysis includes four models:
**unweighted linear models** and **weighted linear models** for both
candidates. These models predict each candidate’s support as a
continuous outcome across different states, using **pollster**,
**national poll**, and **population** as key predictors. The weighted
models additionally incorporate adjustments for recency, sample size,
and poll quality to account for potential biases. This setup aligns with
the New York Times methodology for election polling, which assigns
weights based on poll quality, sample size, and recency to better
reflect voter sentiment [@nyt_polling_averages].

The validation and diagnostic details of the model are provided in [Appendix B](#sec-model-details)

@@

## Recent 3 Months Support Trends Linear Model Prediction

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
set.seed(233)
new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 1000 
  ),
  national_poll = rep(c(0, 1), length(1000)),
  pollster = factor(sample(levels(harris_data$pollster), size = 1000, replace = TRUE), 
                    levels = levels(harris_data$pollster)),  
  population = factor(sample(c("lv","rv"), size = 1000, replace = TRUE)),
  sample_size = sample(1:2300, size = 1000, replace = TRUE),
  pollster_quality= sample(2:4, size = 1000, replace = TRUE)
)|>
  mutate(
    # Sample size weight (assuming a maximum of 2300 responses)
    sample_size_weight = pmin(sample_size / 2300, 1),  # Set sample size; adjust accordingly

    # Pollster quality weight (assuming you want to use a dummy numeric grade, e.g., 2)
    pollster_quality_weight = pollster_quality / 4,

    # Recency weight (this can be based on the date difference from a certain cutoff, e.g., "2024-11-05")
    recency = as.numeric(difftime( as.Date("2024-11-05"),as.Date(end_date), units = "days")),  # Adjust for the sequence
    recency_weight = exp(-recency * 0.1),  # Adjust decay rate as needed
        recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1),
    # Combine all weights into a final combined weight
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight,
    candidate="Harris"
  )
new_data_trump <- data.frame(
  end_date = seq(
    min(trump_data$end_date),
    max(trump_data$end_date),
    length.out = 1000
  ),
  pollster = factor(sample(levels(trump_data$pollster), size = 1000, replace = TRUE), 
                    levels = levels(trump_data$pollster)),  
  population = factor(sample(c("lv","rv"), size = 1000, replace = TRUE)),
   national_poll = rep(c(0, 1), length(1000)),  
  sample_size =sample(1:2300, size = 1000, replace = TRUE),
  pollster_quality= sample(2:4, size = 1000, replace = TRUE)
)|>
  mutate(
    # Sample size weight (assuming a maximum of 2300 responses)
    sample_size_weight = pmin(sample_size / 2300, 1),  # Set sample size; adjust accordingly

    # Pollster quality weight (assuming you want to use a dummy numeric grade, e.g., 2)
    pollster_quality_weight = pollster_quality / 4,

    # Recency weight (this can be based on the date difference from a certain cutoff, e.g., "2024-11-05")
    recency = as.numeric(difftime( as.Date("2024-11-05"),as.Date(end_date), units = "days")),  # Adjust for the sequence
    recency_weight = exp(-recency * 0.1),  # Adjust decay rate as needed

    # Poll frequency weight (you can set a condition based on a hypothetical recent poll count)
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1),

    # Combine all weights into a final combined weight
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight,
    candidate="Trump"
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-support-predicted
#| fig-cap: Predicted Support Trends for Kamala Harris and Donald Trump using Fitted Linear Model
# Predictions for Harris
new_data_harris <- new_data_harris |>
  mutate(
    predicted_unweighted = predict(harris_unweighted_model, newdata = new_data_harris),
    predicted_weighted = predict(harris_weighted_model, newdata = new_data_harris)
  )

# Predictions for Trump
new_data_trump <- new_data_trump |>
  mutate(
    predicted_unweighted = predict(trump_unweighted_model, newdata = new_data_trump),
    predicted_weighted = predict(trump_weighted_model, newdata = new_data_trump)
  )

# Define the cutoff date for recent data
recent_date_cutoff <- as.Date("2024-08-01")

# Filter combined_predictions to include only data from the recent date cutoff onward
combined_predictions<- bind_rows(new_data_harris, new_data_trump)
combined_predictions_filtered<-combined_predictions|>filter(end_date>=recent_date_cutoff)
# Determine the maximum end date for setting the x-axis limit
max_end_date <- max(combined_predictions_filtered$end_date)

# Plotting with the filtered data
ggplot(combined_predictions_filtered, aes(x = end_date, color = candidate)) +
  geom_smooth(aes(y = predicted_unweighted, linetype = "Unweighted"), 
              method = "loess", se = TRUE, size = 1) +  # Unweighted model with smoothing
  geom_smooth(aes(y = predicted_weighted, linetype = "Weighted"), 
              method = "loess", se = TRUE, size = 1) +
  labs(title = "Predicted Support for Kamala Harris and Donald Trump in 2024 US Election \n(From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Weights: recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight\nLinear Regression Model: candidate support percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_linetype_manual(name = "Linear Model Type", values = c("Unweighted" = "dashed", "Weighted" = "solid")) +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

The predicted support for Kamala Harris and Donald Trump in the upcoming 2024 U.S. presidential election is illustrated in Figure 1. This analysis spans the period from August 1, 2024, to October 27, 2024, and uses multiple linear regression models to project voter support.

The un-weighted model predictions indicate that Kamala Harris maintains a consistent support level, with projections hovering around 49%. The blue dashed line represents this trend, showing only minor fluctuations throughout the polling period. Conversely, the weighted model for Harris, indicated by the solid blue line, suggests a slight upward trend in predicted support, reflecting a potential positive shift in voter sentiment as the election date approaches.

In contrast, Donald Trump's predicted support is represented by the red lines. The unweighted model forecasts his support to remain around 48%, while the weighted model predicts a gradual decline to approximately 46%. The red dashed line indicates the unweighted prediction, whereas the solid red line shows the weighted prediction. This downward trajectory may indicate challenges for Trump in sustaining voter enthusiasm amid changing public opinions.

At the conclusion of the polling period, the models suggest that Kamala Harris leads Donald Trump by approximately 4% in the un-weighted (50% vs. 46%) and 2.5% in weighted models (50% vs. 47.5%).

@@

# Additional data details{#first-model-details}

@@

```{r}
#| echo: false
#| message: false
#| warning: false
# Load necessary libraries
library(boot)    # For bootstrapping
library(caret)   # For cross-validation

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Diagnostics of model using residual vs fitted plot and norm Q-Q plot\n-Support for Harris"
#| fig-subcap: ["Residual Plot for Unweighted Model", "Q-Q Plot for Unweighted Model", "Residual Plot for Weighted Model", "Q-Q Plot for Weighted Model"]
#| layout-ncol: 2
